Clouds and Sensitivities Across a Hierarchy of GFDL CMIP6 Models

### 2019-10-14 Silvers_WalkerCell.tex

@@

@@

%\begin{figure}
%  \centering
%      \includegraphics[width=0.6\columnwidth]{walkerfigs/Precip_DMN_4models_1yr.eps}
%  \caption{Domain mean precipitation as a function of time.  All data has been smoothed with a running mean filter.  
%  Thick lines are the control experiments, thin lines have no LW CRE.  All data is shown for the E1 and E2 cases
%  while the P25 and P100 cases show only the first out of five years of the time series.}
%    \label{fig:precip_dom_mn}
%\end{figure}

@@

@@

There is a rich literature on tropical overturning circulations.  Much of it is focused on the Hadley circulation and while some
work has looked at the Walker circulation, it has received less attention.  Part of our motivation in using the framework of the 
Walker circulation comes from the fact that all of the major types of tropical cloud types occur regularly within this circulation
and it thus serves as an ideal testbed/laboratory for studying the interactions and interdependencies of different types of 
clouds and the circulation.  While this study has interpreted the experiments in the context of the Walker Circulation, our results 
are also relevant to the deep overturning circulations and meridional SST gradients that define the ITCZ and the Hadley 
Circulation.  In that context, our results are consistent with those of several recent studies (e.g. Harrop and Hartmann, 2016, 
Popp and Silvers, 2017, Dixit et al. 2018, Flaschner et al. 2018).  These studies, as well as the present one, all show that the 
LWCRE acts to constrain the deep convective region.  This results from an increased atmospheric energy uptake and 
strengthening of the overturning circulation where the deep convective clouds occur (Popp and Silvers, 2017).  Consistent with 
these papers, the present work also shows that the LWCRE has a strong influence on the low-level circulation.   When 
the LWCRE is turned off, the low-level circulations shift upward and are not as well organized (reference psi figures).
There is a corresponding change in the low-level cloud fields, longwave radiative cooling, and the domain mean precipitation.
For the experiments with a GCM-like configuration, the LWCRE strongly influences the precipitation from the large-scale
cloud scheme while leaving the precipitation from the convective parameterization scheme largely unchanged.  This leads 
to a much stronger response of the GCM-like experiments to the LWCRE, especially in the low-levels of the troposphere.  
Because we expect the fraction of precipitation that is due to the convective parameterization to be model dependent 
this could explain why GCMs have a large spread in their response to turning off the LWCRE (reference some 
of Aiko's work?).

>>>>

%One of the most prominent features of our 25km grid-spacing simulations is an asymmetry in both the time-dependent and 
%steady-state solutions.  This results in the steady-state precipitation maximum locating off the maximum in SST and is apparent in 
%the vertical velocity, mass circulation, relative humidity, specific humidity, and radiative heating.   A tendency of the precipitation to 
%occur off of the SST maximum is also apparent in the simulations with a grid-spacing of 100km, although the 25km simulations display 
%an irregularity that is greater than the 100km simulations.  Complex patterns of precipitation over a fixed sinusoidal or Gaussian SST 
%distribution has been noted many times 
%(Bretherton et al. 2006, Kuang and Wofsy, Ming Zhao (personal communication, see Jeevanjee et al. 2018)) but the irregularities have 
%tended to be symmetric, unlike our results.   
%
%For the simulations that are closest to the set up of a GCM, with fully parameterized convection, the steady-state circulation and 
%precipitation maximum are located not over the warmest SST but shifted to regions about one degree cooler.  For the case
%with a grid-spacing of 25km and a domain width of 4000km, the precipitation maximum can remain shifted off of the maximum in
%SST for multiple years.   While focused within the central quarter of the domain, simulations on larger domains and with a 
%grid-spacing of 100km also show an aversion to settling over the SST maximum.   When the convective parameterization
%is turned off, the overturning circulation becomes weaker and broader (w, and psi), but the precipitation, cloud fields, and 
%circulation all consistently reside over the SST maximum.  When the convective parameterization is turned on, a domain mean 
%wind shear develops throughout the lowest 10km that acts to shift the convective activity and circulation off of the SST maximum.  
%Note: check the wind shear for P25large, P100small, and P100large.

@@

Because of the close connection between the domain mean value of precipitation, atmospheric condensational heating, and 
the total radiative cooling the time evolution of the precipitation serves as a useful measure of wether a model has 
reached statistical equilibrium.  Figure \ref{fig:precip_dom_mn} demonstrates that after about 30 days, the P25, E2, and E1
simulations begin to oscillate about a mean state that is roughly constant over periods greater than about 30 days. 
The large oscillations in domain mean precipitation are similar to those noted in previous studies 
(Silvers et al., 2016; Patrizio and Randall, 2019).   The experiments with a higher resolution have a lower value of domain
mean precipitation, with the E2 simulation having the lowest (3.1 mm/day).  This can be understood as a consequence
of the differences in upper level cloud fraction and will be further discussed.

Superposing the circulation and relative humidity illustrates the subsidence driven 
drying and the strong moistening that results from ascent over the region of maximum latent heat flux.

@@

@@

%  \caption{Variables important to boundary layer energy fluxes.  Top left panel shows the surface enthalpy flux ($\rm{W m^{-2}}$) and top right panel shows the equivalent potential temperature on the model level closest to the surface.  The panel on bottom left shows the $u$ wind ($\rm{m s^{-1)}}$)
%  on the model level closest to the surface and the panel on bottom right shows the domain mean
%  $\rm{u}$ wind.}

@@

There is a rich literature on tropical overturning circulations.  Much of it is focused on the Hadley circulation and while some
work has looked at the Walker circulation, it has received less attention.  Part of our motivation in using the framework of the 
Walker circulation comes from the fact that all of the major types of tropical cloud types occur regularly within this circulation
clouds and the circulation.  While this study has interpreted the experiments in the context of the Walker Circulation, our results 
are also relevant to the deep overturning circulations and meridional SST gradients that define the ITCZ and the Hadley 
Circulation.  In that context, our results are consistent with those of several recent studies (e.g. Harrop and Hartmann, 2016, 
Popp and Silvers, 2017, Dixit et al. 2018, Flaschner et al. 2018).  These studies, as well as the present one, all show that the 
LWCRE acts to constrain the deep convective region.  This results from an increased atmospheric energy uptake and 
strengthening of the overturning circulation where the deep convective clouds occur (Popp and Silvers, 2017).  Consistent with 
these papers, the present work also shows that the LWCRE has a strong influence on the low-level circulation.   When 
the LWCRE is turned off, the low-level circulations shift upward and are not as well organized (reference psi figures).
There is a corresponding change in the low-level cloud fields, longwave radiative cooling, and the domain mean precipitation.
For the experiments with a GCM-like configuration, the LWCRE strongly influences the precipitation from the large-scale
cloud scheme while leaving the precipitation from the convective parameterization scheme largely unchanged.  This leads 
to a much stronger response of the GCM-like experiments to the LWCRE, especially in the low-levels of the troposphere.  
Because we expect the fraction of precipitation that is due to the convective parameterization to be model dependent 
this could explain why GCMs have a large spread in their response to turning off the LWCRE (reference some 
of Aiko's work?).

### 2019-10-14 Silvers_cfmip_draft.tex

During the developmental of the AM4.0 GFDL model particular care was given to ensuring that the energy budget at 
the top of the atmosphere compared well with observations.  The data set that was most heavily leaned on to ensure 
this was the CERES-EBAF TOA radiative fluxes, version 2.8.  Based on AM4.0 data that was used for the CMIP6 AMIP
experiment, the total TOA flux bias is 989898, the bias in CRE 
due to longwave radiative fluxes is $-2.4 \rm{W m^{-2}} $ and the bias due to CRE of the shortwave radiative fluxes   
is $-1.1 \rm{W m^{-2}} $.  About the same time the AM4.0 development was frozen and the process of documentation 
underway, the CERES-EBAF was updated to version 4.1.  When comparing the same AM4.0 AMIP experiment to the 
newer version of CERES-EBAF data we find biases of $-2.2 \rm{W m^{-2}} $ for the longwave CRE and 
$-2.9 \rm{W m^{-2}} $ for the shortwave CRE.   Should these biases be computed using only the AMIP years that 
correspond to the CERES years?   As previously discussed (e.g. Kay et al., 2012) these longwave and shortwave CRE 
biases are difficult to interpret because model developers often tune models to match particular CERES-EBAF data 
sets despite the well known fact that different versions of the CERES-EBAF data lead to changes in the observed CRE of
several $\rm{W m^{-2}}$.   Our aim in discussing this for AM4.0 is transparency in the model development process, as 
well as to point out that both versions of the CERES-EBAF data show similar shortcomings of the spatial patters of the 
simulated clouds  from AM4.0 in the AMIP experiment.

@@

In an attempt to compare to the values of feedback that were reported in Ringer et al., 2014, we use the same method 
to compute the radiative feedbacks for the GFDL models.  That is, for the abrupt $4xCO_{2}$ experiment, feedbacks are 
inferred to be the slopes of a linear regression calculation using annual, global-mean top-of-atmosphere flux anomalies (Gregory et al., 2004).  From the atmosphere only experiments (amip-p4K, amipFuture, and aqua-p4K), the feedbacks are calculated as 
differences in top-of-atmosphere flux anomalies between the perturbation and control experiments.

@@

\begin{table}
\begin{center}
\caption{Global mean radiative feedbacks.  Compare to Ringer et al. 2014.}
    \begin{tabular}{*{5}{c}}
    \hline
    \hline
 Feedback & Aqua p4K & Amip p4K & Amip Future & Abrupt 4xCO2    \\ \hline
    Net          &   -2.2    &  -1.6              & -1.8           &    -0.84 (from Tim)         \\ 
    \\
    Net CRE      & 0.3      & -0.1            & 0.1           &?            \\  
    \\
    LW CRE       & 0.4       & 0.1            & 0.1           & ?              \\  
    \\
    SW CRE      & -0.7       & -0.1          & -0.2          & ?              \\  
    \\
    LW CLR       & 2.0       & 2.0             & 2.0           & ?             \\  
    \\
    SW CLR      & -0.1      & -0.3            & -0.3          & ?                   \\  \hline

@@

>>>>

During the developmental of the AM4.0 GFDL model particular care was given to ensuring that the energy budget at 
the top of the atmosphere compared well with observations.  The data set that was most heavily leaned on to ensure 
this was the CERES-EBAF TOA radiative fluxes, version 2.8 (Loeb et al., 2009).  Based on AM4.0 data that was used for 
the CMIP6 AMIP
experiment, the total TOA flux bias is 989898, the bias in CRE 
due to longwave radiative fluxes is $-2.4 \rm{W m^{-2}} $ and the bias due to CRE of the shortwave radiative fluxes   
is $-1.1 \rm{W m^{-2}} $.  About the same time the AM4.0 development was frozen and the process of documentation 
underway, the CERES-EBAF was updated to version 4.1.  When comparing the same AM4.0 AMIP experiment to the 
newer version of CERES-EBAF data we find biases of $-2.2 \rm{W m^{-2}} $ for the longwave CRE and 
$-2.9 \rm{W m^{-2}} $ for the shortwave CRE.   Should these biases be computed using only the AMIP years that 
correspond to the CERES years?   As previously discussed (e.g. Kay et al., 2012) these longwave and shortwave CRE 
biases are difficult to interpret because model developers often tune models to match particular CERES-EBAF data 
sets despite the well known fact that different versions of the CERES-EBAF data lead to changes in the observed CRE of
several $\rm{W m^{-2}}$.   Our aim in discussing this for AM4.0 is transparency in the model development process, as 
well as to point out that both versions of the CERES-EBAF data show similar shortcomings of the spatial patters of the 
simulated clouds  from AM4.0 in the AMIP experiment.

Paragraph summarizing the basic characteristics of clouds in AM4.0 as stated by Ming's papers.  We could also summarize 
what David finds about clouds in AM4.0 in his paper.

@@

In order to compare to the values of feedback that were reported in Ringer et al., 2014, we use the same method 
to compute the radiative feedbacks for the GFDL models.  That is, for the abrupt $4xCO_{2}$ experiment, feedbacks are 
inferred to be the slopes of a linear regression calculation using annual, global-mean top-of-atmosphere flux anomalies 
(Gregory et al., 2004).  From the atmosphere only experiments (amip-p4K, amipFuture, and aqua-p4K), the feedbacks are calculated as 
differences in top-of-atmosphere flux anomalies between the perturbation and control experiments.

@@

\begin{table}
\begin{center}
\caption{Global mean radiative feedbacks.  The values for the Abrupt 4xCO2 experiment
have been estimated from a linear fit of the years 51-300.  Compare to Ringer et al. 2014.}
    \begin{tabular}{*{5}{c}}
    \hline
    \hline
 Feedback & Aqua p4K & Amip p4K & Amip Future & Abrupt 4xCO2    \\ \hline
    Net          &   -2.2    &  -1.6              & -1.8           &    -0.46 (-0.84; Tim)         \\ 
    \\
    Net CRE      & 0.3      & -0.1            & 0.1           & 0.65   \\  
    \\
    LW CRE       & 0.4       & 0.1            & 0.1           & 0.14    \\  
    \\
    SW CRE      & -0.7       & -0.1          & -0.2          & -0.5              \\  
    \\
    LW CLR       & 2.0       & 2.0             & 2.0           & 1.7             \\  
    \\
    SW CLR      & -0.1      & -0.3            & -0.3          & -0.6                   \\  \hline

@@

\section{What are we learning?}

Experiments using Cess-like uniform SST warming as a proxy for climate change have successfully predicted that much of the differences 
among models is due to the cloud radiative affect.  However, we also know that the sensitivity derived from Cess-like experiments is not
a good predictor of the climate sensitivity for a parent, coupled model.  Comparison of Cess-sensitivities and coupled model sensitivities 
combined with patterned prescribed SST experiments using AGCM have demonstrated the significant role of a pattern-effect in determining 
the sensitivity of a model.

We also know that a method which uses linear regression of a constant forcing experiment to estimate the climate sensitivity only 
provides an accurate measure of the climate sensitivity when the initial 20-50 years of simulation are excluded from the analysis.  
The climate sensitivities of models that have been run to equilibrium with millenial-scale simulations show a large discrepency in the
different time periods used to estimate this sensitivity.

Clouds in a climate model are highly maleable while also being a dominant factor controlling what the climate sensitivity is for that model.  
Yet we do not have satisfactory clouds constraints to impose on the clouds of a climate model.  If improving the representation of low-level
clouds in the GFDL model was the first priority, it could be done.  However it is not the first priority and doing so degrades other elements of 
the climate.
